---
title: "CRISPRAid-tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CRISPRAid-tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

`CRISPRAid` is an R package designed to streamline CRISPR-Cas9 data analysis, focusing on tasks such as guide RNA (gRNA) efficiency prediction, off-target analysis, and indel detection. This vignette provides an overview of the package’s main functions and includes examples for each.

To get started, load the package:

```{r setup}
library(CRISPRAid)
```

# Analyzing gRNA Efficiency

The analyze_gRNA_efficiency() function predicts the editing efficiency of a given gRNA sequence against a target DNA sequence. Here’s how to use it:

```{r analyze_gRNA_efficiency_example}
# Example gRNA and target DNA sequences
gRNA_seq <- "GACGTCTAGT"
target_seq <- "TGCTACGTAGT"

# Analyze efficiency
efficiency_result <- analyze_gRNA_efficiency(gRNA_seq, target_seq)
print(efficiency_result)
```

# Visualizing gRNA Efficiency

The visualize_gRNA_efficiency() function creates a heatmap for comparing gRNA efficiency across multiple sequences.

```{r visualize_gRNA_efficiency_example}
# Example data for multiple gRNA efficiencies
efficiency_data <- data.frame(
  gRNA = c("gRNA1", "gRNA2", "gRNA3"),
  efficiency = c(80, 90, 75)
)

# Generate heatmap of gRNA efficiencies
efficiency_plot <- visualize_gRNA_efficiency(efficiency_data)
print(efficiency_plot)
```

# Evaluating gRNA Specificity

The evaluate_gRNA_specificity() function calculates the specificity of a gRNA sequence by comparing it with a set of potential off-target sequences. This can help in identifying sequences that may lead to unintended edits.

```{r evaluate_gRNA_specificity_example}
# Example gRNA sequence and potential off-target sequences
gRNA_seq <- "GACGTCTAGT"
off_target_seqs <- c("GACGTCTTGT", "TGCATCTAGG", "GACGTGTAGT")

# Evaluate specificity
specificity_result <- evaluate_gRNA_specificity(gRNA_seq, off_target_seqs)
print(specificity_result)
```

# Simulating Off-Target Effects in a Genome

The simulate_off_target_effects() function simulates potential off-target binding sites across a genome. This is useful for preliminary analysis to estimate the likelihood of unintended edits.

```{r simulate_off_target_effects_example}
# Simulate off-target effects in a genome of length 1000 with 10 potential sites
gRNA_seq <- "GACGTCTAGT"
genome_length <- 1000
num_sites <- 10

off_target_simulation <- simulate_off_target_effects(gRNA_seq, genome_length, num_sites)
print(off_target_simulation)
```

# Plotting Off-Target Effects

The plot_off_target_map() function allows users to visualize off-target effects on the genome using an interactive plot. This is useful for identifying potential off-target binding sites.

```{r plot_off_target_map_example}
# Example data for off-target effects
off_target_data <- data.frame(
  position = seq(1, 100, by = 10),
  effect_size = runif(10, min = 0, max = 1)
)

# Plot interactive map of off-target effects
off_target_plot <- plot_off_target_map(off_target_data)
off_target_plot
```

# Visualizing Base Composition of gRNA Sequences

The visualize_gRNA_base_composition() function generates a bar plot showing the base composition (A, T, G, C) of a set of gRNA sequences. This can provide insight into the stability and characteristics of the sequences.

```{r visualize_gRNA_base_composition_example}
# Example set of gRNA sequences
gRNA_seqs <- c("GACGTCTAGT", "TGCATCTAGG", "GACGTGTAGT")

# Generate base composition plot
base_composition_plot <- visualize_gRNA_base_composition(gRNA_seqs)
print(base_composition_plot)
```

# Computing Indels

The compute_indels() function calculates indel frequencies from sequencing data, based on a user-defined threshold.

```{r compute_indels_example}
# Example sequencing data with mutation rates
seq_data <- data.frame(
  mutation_rate = c(0.05, 0.2, 0.15, 0.01, 0.3)
)

# Compute indels with a threshold of 0.1
indel_result <- compute_indels(seq_data, indel_thres = 0.1)
print(indel_result)
```

# Conclusion

This vignette provides an overview of the main functions in CRISPRAid package:
- gRNA efficiency prediction (analyze_gRNA_efficiency)
- gRNA efficiency visualization (visualize_gRNA_efficiency)
- specificity analysis (evaluate_gRNA_specificity)
- off-target effect simulation (simulate_off_target_effects)
- off-target effect visualization (plot_off_target_map)
- base composition visualization (visualize_gRNA_base_composition)
- indel frequency computation (compute_indels)




